Estimation of Stochastic Wilson and Cowan Neural Field Model using Expectation Maximisation Algorithm

### 2011-07-16 ltx/data_paper/DataPaper.tex
edits to discussion

The estimated supports were then employed in the EM algorithm described in Section \ref{sec:EstimationAlgorithm} to estimate the smoothed states $\hat{\mathbf x}_t^b$, the connectivity kernel weights $\boldsymbol\theta$, the synaptic dynamic $\xi$, and the hyperparameters $\sigma_{\epsilon}^2$ and $\sigma^2_d$. A  square grid of $n_{\theta}=25$ (5 by 5) basis functions with $2.25~mm$ spacing were used to cover the estimated kernel support where $\sigma_{\psi}^2=3.5$. This is shown in Figure.\ref{fig:KernelWidthEstimation} by blue circles, each of which representing full width at half maximum of the connectivity kernel basis functions. The true and estimated connectivity kernel for 100 Monte Carlo simulations are shown in Figure~\ref{fig:KernelEstimates}. Cross-sections of the connectivity kernel estimates for each run are shown in Figure~\ref{fig:KernelEstimateCrossSection}. The reconstructed kernel is in good accordance with the actual kernel, where the actual kernel is inside the confidence interval. The large standard deviation is due to the high number of parameters need to be estimated. Figure~\ref{fig:Histogram} depicts the histogram of parameter estimates from 100 simulations where the true parameter value is shown by the solid line and the mean parameter estimate is shown by the dashed line. These histograms indicate small bias in parameter estimates with a very small standard deviation. Examples of neural field reconstruction are given in Figure.~\ref{fig:FieldEstimates}.  The figure shows snapshots at three time instants of the true neural field and the estimated field. In each subplot neural activity is plotted against spatial index in 1D to provide a comparision between true and the estimated underlying fields. Plot of lower bound of the log likelihood versus iterations of the EM is shown in Figure.~\ref{fig:LowerBound}. In general, for the non-linear dynamical systems the actual lower bound can not be claculated analytically, therefore the approximate of the lower bound obtained by URTSS is shown. The change in lower bound drops below 0.009 percent after 15 iterations, confirming
that the algorithm converged to steady parameter values.
\section{Discussion}
The correlation analysis specifies approximate supports for both disturbance covariance function and connectivity kernel. Since the analysis assumes the sigmoidal firing rate to be in the active region, the amplitudes for the kernel and the disturbance cannot be inferred this way. Although the unknown parameters $\xi$ and $\sigma_{\varepsilon}^2$ can be chosen as described in Correlation analysis section, their values influence the exact shape of the kernel significantly, this step might be considered as an initialisation for the EM algorithm. Knowing the spatial extent of the kernel greatly increases the speed of the algorithm as the number of basis functions to decompose the kernel reduces significanly compared to the case in which the basis funtions are laid out across the entire spatial domain. This also improves the uncertainty of the parameter estimates as the number of parameters to be estimated reduces. On the other hand, the approximated support of the disturbance covariance function facilitates the implementation of the EM algorithm as the disturbance covariance matrix can be factorised into a constant matrix which only depends on the inferred support and an unknown scaling parameter which needs to be estimated.

>>>>

The estimated supports were then employed in the EM algorithm described in Section \ref{sec:EstimationAlgorithm} to estimate the smoothed states $\hat{\mathbf x}_t^b$, the connectivity kernel weights $\boldsymbol\theta$, the synaptic dynamic $\xi$, and the hyperparameters $\sigma_{\epsilon}^2$ and $\sigma^2_d$. A  square grid of $n_{\theta}=25$ (5 by 5) basis functions with $2.25~mm$ spacing were used to cover the estimated kernel support where $\sigma_{\psi}^2=3.5$. This is shown in Figure.\ref{fig:KernelWidthEstimation} by blue circles, each of which representing full width at half maximum of the connectivity kernel basis functions. The true and estimated connectivity kernel for 100 Monte Carlo simulations are shown in Figure~\ref{fig:KernelEstimates}. Cross-sections of the connectivity kernel estimates for each run are shown in Figure~\ref{fig:KernelEstimateCrossSection}. The reconstructed kernel is in good accordance with the actual kernel, where the actual kernel is inside the confidence interval. The large standard deviation is due to the high number of parameters need to be estimated. Figure~\ref{fig:Histogram} depicts the histogram of parameter estimates from 100 simulations where the true parameter value is shown by the solid line and the mean parameter estimate is shown by the dashed line. These histograms indicate small bias in parameter estimates with a very small standard deviation. Examples of neural field reconstruction are given in Figure.~\ref{fig:FieldEstimates}.  The figure shows snapshots at three time instants of the true neural field and the estimated field. In each subplot neural activity is plotted against spatial index in 1D to provide a comparision between true and the estimated underlying fields. Plot of lower bound of the log likelihood versus iterations of the EM is shown in Figure.~\ref{fig:LowerBound}. In general, for the non-linear dynamical systems the actual lower bound can not be claculated analytically, therefore the approximate of the lower bound obtained by URTSS is shown. The change in lower bound drops below 0.009 percent after 15 iterations, confirming
that the algorithm converged to steady parameter values.
\section{Discussion}
The correlation analysis specifies approximate supports for both disturbance covariance function and connectivity kernel. Although the unknown parameters $\xi$ and $\sigma_{\varepsilon}^2$ can be set as described in \S~\ref{subsec:CorrAnal}, the chosen values influence the exact shape of the kernel significantly, the analysis also assumes the sigmoidal firing rate to be in the active region, as a result  the kernel and the disturbance amplitudes cannot be inferred only through correlation analysis. This step might be considered as a prerequisite for the EM algorithm described in \S~\ref{sec:EstimationAlgorithm} where approximated supports are employed to estimate the gain of the kernel and disturbance. Knowing the spatial extent of the kernel greatly increases the speed of the algorithm as the number of basis functions to decompose the kernel reduces significanly compared to the case in which the basis funtions are laid out across the entire spatial domain. This also improves the uncertainty of the estimates because of the lower number of parameters needs to be estimated. On the other hand, the approximated support of the disturbance covariance function facilitates the implementation of the EM algorithm as the disturbance covariance matrix can be factorised into two components, a constant matrix depending only on the inferred support and an unknown scaling parameter. This way the problem of estimating the covariance matrix reduces to estimation of a single scaling parameter.
